---
slug: "dividend-yield-tr"
date: "2020-04-05"
title: "Calculation of dividend yield from index returns"
tags: ["R", "tidyquand", "finance", "stock market"]
authors: [mgei]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, warning = F, message = F)
```

```{r setupcache, include=FALSE}
knitr::opts_chunk$set(cache = T)
```

## From TR to dividends

If you have a price index $I$ and the corresponding total return index $I_{TR}$, then you can calculate a pre-dividend version of the total return index for period *t* as follows

$$ I^{predividend}_{TR, t} = I_{TR, t-1}*(1+r_{t}) $$

Our pre-dividend total return index of period *t* is the previous period's total return index times the return of the price index. The return is simple arithmetic.

$$ r_{t} = \dfrac{I_{t}}{I_{t-1}} -1 $$

The dividend on the total return index in period *t* $D_{t}$ would be the difference between the actual total return index value and the computed pre-dividend version of it.

$$ D_{t} = I_{TR, t} - I^{predividend}_{TR,t} $$

## Dividend yield

The yield is the computed dividend divided by the total return index. Note that periods should be in years, as the yield is expressed in annual terms. If your *t* is not years, you would need to aggregate (yearly sum of dividends, and mean of TR index value).

$$ y_{dividend} = D_t / I_{TR,t}$$

In R and using the tidyverse and the common finance package tidyquant to get S&P500 data from Yahoo, this is the script that I came up with.

```{r dividend, eval = F}

library(tidyverse)
library(tidyquant)

sp500 <- tq_get("^GSPC", from = "1990-01-01")
sp500_tr <- tq_get("^SP500TR", from = "1990-01-01")

spx <- left_join(sp500 %>% select(date, SP500 = adjusted), 
                 sp500_tr %>% select(date, SP500_TR = adjusted), 
                 by = "date")

spx_dividend <- spx %>% 
  mutate(SP500_return = SP500/lag(SP500)-1,
         SP500_TR_predividend = lag(SP500_TR)*(1+SP500_return),
         dividend = SP500_TR - SP500_TR_predividend)

spx_dividend_yield <- spx %>% 
  group_by(year = year(date)) %>% 
  summarise(dividend = sum(dividend, na.rm = T),
            SP500_TR = mean(SP500_TR)) %>% 
  mutate(dividendyield = dividend/SP500_TR)
```

Posted on quant.stackexchange.com https://quant.stackexchange.com/a/53069/46111