---
slug: "mock-addresses"
date: "2019-01-14"
title: "Generate mock addresses with realistic names and locations"
tags: ["R", "spatial", "mock data", "Youtube", "leaflet", "maps"]
authors: [mgei]
---

> Originally published on a Github gist: https://gist.github.com/mgei/0269fb9e4120be24af6b5d1d344b89b6

# Fake name generator

[Fake Name Generator](https://www.fakenamegenerator.com/order.php) allows to generate mock names/addresses. One can choose from plenty of fields to include besides names and addresses, such as date of birth, phone number, ID number, and also lon/lat coordinates.

For the following purpose we will order as many addresses as possible and should not forget to include longitude and latitude. Once this is done, we get an e-mail notification with the download link. Do not worry, it is free. It is trustworthy, I have not got any newsletter spam after ordering.

# R for subsetting by location

In the next step we will load the data into R. This will allow us to subset the entries in a realistic way. Imagine we are locally operating a company, a hospital or similar. Our clients or patients are likely to come from nearby. The further a person lives, the less likely he is to be your customer. That's what we want to simulate.


```{r setup, warning=F, message=F}
library(tidyverse)
library(lubridate)
library(leaflet)
```


```{r data}

addresses <- read_csv("postdata/fakeaddresses/FakeNameGenerator.com_f201506b.csv")
  
addresses <- addresses %>% mutate(Birthday = mdy(Birthday),
                                  Title = case_when(Title == "Mr." ~ "Herr",
                                  Title == "Mrs." ~ "Frau",
                                  Title == "Ms." ~ "Frau",
                                  Title == "Dr." & Gender == "female" ~ "Frau Dr.",
                                  Title == "Dr." & Gender == "male" ~ "Herr Dr.",
                                  T ~ Title))

head(addresses)

```

Next we define our own location. Around where we are we want the most addresses. Therefore, we add a column *distance*.

```{r loc}
mylat <- 47.5596
mylon <- 7.5886

addresses <- addresses %>% 
  mutate(distance = sqrt((Longitude-mylon)^2+(Latitude-mylat)^2))

```

Finally, we use dplyr's `sample_n()` to subsample the addresses in an inverse exponential way and plot it in a leaflet map.

```{r subsample}

patients <- addresses %>% sample_n(100, weight = 1/distance^2)

patients %>% 
  leaflet() %>% 
  setView(lng = mylon, lat = mylat, zoom = 10) %>% 
  addTiles() %>% 
  addCircleMarkers(label = ~str_c(GivenName, Surname, StreetAddress, City, sep = " "),
                   radius = 2, color = "red") %>% 
  addMarkers(label = "Me", lng = ~mylon, lat = ~mylat)

```

# Youtube tutorial

```{r, eval=TRUE, echo=F}
blogdown::shortcode("youtube", "AGb5Yk_d8Lk")
```
