---
slug: "fuel-consumption"
date: "2019-06-12"
title: "Fuel consumption by model and manufacturer"
tags: ["R", "dplyr", "tabulizer", "plotting"]
authors: [mgei]
---

# European car fuel consumption data

We found data from the Swiss automobile association *TCS*. While they provide a simple online tool [Verbrauchskatelog](http://www.verbrauchskatalog.ch/de/), the data as a whole is downloadable only in PDF format [here](http://www.verbrauchskatalog.ch/?async=1&merge=1).

One can read tables from PDF documents in R using the [tabulizer](https://cran.r-project.org/web/packages/tabulizer/vignettes/tabulizer.html) package. A great example for it's use was published last year on R Views [Player Data for the 2018 FIFA World Cup](https://rviews.rstudio.com/2018/06/14/player-data-for-the-2018-fifa-world-cup/).

In my case, I struggled to even install *tabulizer* on Ubuntu, so I had revert to Windows.

For reading table `extract_tables` is used. Note we set the output to *matrix* as it will be easier to turn it into a tibble or data-frame later.

```{r tabu, eval=F}
library(tabulizer)
library(tidyverse)

out1 <- extract_tables("vkatalog.pdf", output = "matrix", encoding = "UTF-8")
```

Clean the matrix...

```{r process, eval = F}
m <- matrix(ncol = 12)
for (i in 1:(length(out1)-1)) {
  print(i)
  m <- rbind(m, out1[[i]][,1:12])
}
```

...and turn it into a tibble:

```{r tibblemaking, eval = F}
carst <- m %>% as_tibble()

colnames(carst) <- c("Modell", "Neupreis", "TreibstoffCode", "Antrieb", "Getriebe", "Hubraum", 
                     "PS", "Normverbrauch", "Benzinaeq", "CO2", "CO2 Energieerz", "Effizientkat")

# carst %>% saveRDS("data/carst.RDS")
```

From here on it's a data handling task. We shall clean it and make it tidy.

```{r read, echo=F, warning=F, message=F}
library(tidyverse)

carst <- readRDS("postdata/fuel-consumption/carst.RDS")

```


```{r cars, warning=F}
# carst <- carst %>% 
#   filter(!is.na(Modell), Modell != "") %>% 
#   mutate(Hersteller = if_else(Neupreis == "", Modell, NA_character_)) %>% 
#   select(Hersteller, everything()) %>% 
#   fill(Hersteller) %>% 
#   filter(Neupreis != "")

carst_clean <- carst %>% 
  separate(Normverbrauch, into = c("Normverbrauch", "Normverbrauch E"), sep = " \\+ ",
           fill = "right") %>% 
  mutate(Neupreis = str_remove(Neupreis, "'") %>% as.integer(),
         Hubraum = str_remove(Hubraum, "'") %>% as.integer(),
         Normverbrauch = str_remove(Normverbrauch, " l") %>% as.double(),
         `Normverbrauch E` = str_remove(`Normverbrauch E`, " kWh") %>% as.double(),
         PS = as.integer(PS),
         Benzinaeq = str_remove(Benzinaeq, "-") %>% str_trim()) %>%
  rename(`Normverbrauch (l)` = Normverbrauch,
         `Normverbrauch E (kWh)` = `Normverbrauch E`) %>% 
  mutate(Treibstoff = recode(TreibstoffCode, "B" = "Benzin", "D" = "Diesel", "C" = "Hybrid Benzin",
                             "Y" = "Wasserstoff / Elektro", "F" = "Hybrid Diesel", "N" = "Erdgas (CNG)",
                             "E" = "Elektro", R = "Elektro mit Range Extender", "X" = "Wasserstoff / Elektro"),
         Antrieb = recode(AntriebCode, "V" = "Frontantrieb", "H" = "Heckantrieb", "A" = "Allrad"),
         CO2 = if_else(str_detect(Benzinaeq, " ") & CO2 == "", str_sub(Benzinaeq, 4, -1), CO2),
         Benzinaeq = if_else(str_detect(Benzinaeq, " "), str_sub(Benzinaeq, 1, 3), Benzinaeq) %>% as.double(),
         `CO2 klimarel` = if_else(str_detect(CO2, "\\("), str_extract_all(CO2, "\\([^()]+\\)", "") %>% 
                                    str_remove_all("\\(|\\)"), "") %>% as.integer(),
         `CO2 Energieerz` = as.integer(`CO2 Energieerz`),
         CO2 = str_sub(CO2, 1, 3) %>% as.integer()) %>% 
  # rowwise() %>% 
  # mutate(CO2 = str_sub(CO2, 1, 3) %>% as.integer()) %>% 
  # ungroup() %>% 
  select(Hersteller, Modell, Neupreis, TreibstoffCode, Treibstoff, AntriebCode, Antrieb, Getriebe, Hubraum, 
         PS, `Normverbrauch (l)`, `Normverbrauch E (kWh)`, Benzinaeq, CO2, `CO2 klimarel`, `CO2 Energieerz`, Effizientkat)


carst_clean %>% summary()
```


# Graphs

```{r graphs1, warning=F}
carst_clean %>% 
  filter(Hersteller %in% c("Mercedes-Benz", "Audi", "BMW", "Opel", "VW", "Porsche", "Smart"), PS < 500, Treibstoff == "Benzin") %>% 
  ggplot(aes(x = PS, y = `Normverbrauch (l)`)) + 
  geom_point(alpha = 0.3, aes(color = Hersteller)) +
  geom_smooth(method = "loess", formula = y ~ x + x^2, se = F, fullrange = T) +
  theme(legend.position = "bottom") +
  labs(title = "Fahrzeuge deutscher Hersteller",
       subtitle = "Treibstoff: Benzin") +
  scale_x_continuous(limits = c(50, 500)) +
  scale_y_continuous(limits = c(3, 15))
```

```{r graphs2, warning=F}
carst_clean %>% 
  filter(Hersteller %in% c("Peugeot", "DS Automobiles", "Renault", "Alpine", "Citroen"), PS < 500, Treibstoff == "Benzin") %>% 
  ggplot(aes(x = PS, y = `Normverbrauch (l)`)) + 
  geom_point(alpha = 0.3, aes(color = Hersteller)) +
  geom_smooth(method = "loess", formula = y ~ x + x^2, se = F, fullrange = T) +
  theme(legend.position = "bottom") +
  labs(title = "Fahrzeuge französischer Hersteller",
       subtitle = "Treibstoff: Benzin") +
  scale_x_continuous(limits = c(50, 500)) +
  scale_y_continuous(limits = c(3, 15))
```

```{r graphs3, warning=F}
carst_clean %>% 
  filter(Treibstoff %in% c("Benzin", "Diesel")) %>%
  ggplot(aes(x = PS, y = `Normverbrauch (l)`, color = Treibstoff)) + 
  geom_point(alpha = 0.3) +
  geom_smooth(method = "loess", formula = y ~ x + x^2, se = F, fullrange = T) +
  theme(legend.position = "bottom") +
  labs(title = "Verbrauch nach Leistung, Diesel und Benzin") +
  scale_x_continuous(limits = c(50, 500)) +
  scale_y_continuous(limits = c(3, 15))
```

```{r graphs4, warning=F, fig.height=10}
carst_clean %>%
  filter(!is.na(CO2)) %>% 
  group_by(Hersteller) %>% 
  filter(n() > 10) %>%
  mutate(meanCO2 = mean(CO2)) %>% 
  ungroup() %>%
  ggplot(aes(x = reorder(Hersteller, desc(meanCO2)), y = CO2)) +
  geom_point(aes(color = Hersteller), alpha = 0.3) +
  geom_boxplot(alpha = 0.1, outlier.alpha = 0) +
  coord_flip() +
  labs(title = "CO2 Emission nach Hersteller/Modelle", x = "") +
  theme(legend.position = "none")
  
```

