---
slug: "pca-usecase"
date: "2019-04-23"
title: "PCA use case"
tags: ["PCA", "R", "tidyquant", "finance", "stock market", "linear model"]
authors: [mgei]
---

> Originally published on RPubs: http://rpubs.com/mgei/pca-usecase

```{r setup, message=F, warning=F}

library(tidyverse)
library(tidyquant)
library(quantmod)
library(lubridate)
library(broom)      # for tidying up linear models
library(factoextra)

```

```{r setupcache, include=FALSE}
knitr::opts_chunk$set(cache = T)
```

# Data

FTSE 350 Industrial Engineering [index](https://www.investing.com/indices/industrial-engineering-historical-data). The index components are Bodycote, Hill&Smith, IMI, Rotork, Spirax-Sarco Engineering, and Weir Group ([components](https://www.investing.com/indices/industrial-engineering-components)).

```{r indexdata, eval = F}

index <- read_csv("postdata/pca-usecase/FTSE 350 Industrial Engineering Historical Data.csv")

index %<>%
  rename(date = Date) %>% 
  mutate(date = mdy(date),
         adjusted = Price) %>% 
  select(-Vol., -`Change %`)

# gets data from Yahoo Finance
bodycote   <- tq_get("BOY.L")
hill_smith <- tq_get("HILS.L")
imi        <- tq_get("IMI.L")
rotork     <- tq_get("ROR.L")
spirax_s   <- tq_get("SPX.L")
weir       <- tq_get("WEIR.L")

data <- bind_rows(index %>%      mutate(type = "index", name = "FTSE 350 Industrial Engineering"), 
                  bodycote %>%   mutate(type = "stock", name = "Bodycote plc"), 
                  hill_smith %>% mutate(type = "stock", name = "Hill & Smith Holdings PLC"), 
                  imi %>%        mutate(type = "stock", name = "IMI plc"), 
                  rotork %>%     mutate(type = "stock", name = "Rotork plc"), 
                  spirax_s %>%   mutate(type = "stock", name = "Spirax-Sarco Engineering plc"), 
                  weir %>%       mutate(type = "stock", name = "The Weir Group PLC"))

```

```{r readdata, echo=F, warning=F, message=F}

data <- readRDS("postdata/pca-usecase/data.RDS")

```

Compute the returns.

```{r retcomp, warning=F}

# compute returns (daily)
data_rets <- data %>% 
  select(date, name, type, adjusted) %>% 
  arrange(type, name, date) %>% 
  group_by(name, type) %>%
  mutate(return = adjusted/lag(adjusted)-1,
         return_ln = log(adjusted/lag(adjusted))) %>% 
  filter(!is.na(return_ln),
         # cut-off date
         date >= "2016-04-15") %>% 
  mutate(cumret_ln = cumsum(return_ln),
         adjusted_ln = exp(cumret_ln))

data_rets %>% 
  ggplot(aes(x = date, y = adjusted_ln, color = name)) +
  geom_line(aes(size = factor(desc(type)))) +
  scale_size_discrete(range = c(0.4,1.4), guide = "none") +
  theme_tq()

# compute returns (monthly)
data_rets_m <- data %>% 
  select(date, name, type, adjusted) %>% 
  arrange(type, name, date) %>% 
  group_by(name, type, date = floor_date(date, unit = "months") %>% date()) %>%
  summarise(adjusted = first(adjusted)) %>% 
  mutate(return = adjusted/lag(adjusted)-1,
         return_ln = log(adjusted/lag(adjusted))) %>% 
  filter(!is.na(return_ln),
         # cut-off date
         date >= "2016-04-01") %>% 
  mutate(cumret_ln = cumsum(return_ln),
         adjusted_ln = exp(cumret_ln))

data_rets_m %>% 
  ggplot(aes(x = date, y = adjusted_ln, color = name)) +
  geom_line(aes(size = factor(desc(type)))) +
  scale_size_discrete(range = c(0.4,1.4), guide = "none") +
  theme_tq()

```

Spread format for modelling.

```{r spreadformat}

data_rets %>%
  ungroup() %>% 
  select(date, name, return_ln) %>% 
  filter(date >= "2016-04-15") %>% 
  spread(name, return_ln) -> returns_daily

data_rets_m %>%
  ungroup() %>% 
  select(date, name, return_ln) %>% 
  filter(date >= "2016-04-01") %>% 
  spread(name, return_ln) -> returns_monthly

```

Linear model. The *estimates* can be seen as weights.

```{r linmodel}

lm(`FTSE 350 Industrial Engineering` ~ . -1, data = returns_daily %>% 
     select(-date)) -> model

tidy(model)

```

Estimated/weights shoud approximately sum up to 1.

```{r sumofweights}

tidy(model) %>% 
  summarise(total = sum(estimate))

# note R-square of .99
glance(model)$r.squared

```

Correlation plots, daily and monthly.

```{r corr}

data_rets %>% 
  ungroup() %>% 
  select(date, name, return_ln) %>% 
  spread(name, return_ln) %>% 
  select(-date) %>% 
  chart.Correlation()

data_rets_m %>% 
  ungroup() %>% 
  select(date, name, return_ln) %>% 
  spread(name, return_ln) %>% 
  select(-date) %>% 
  chart.Correlation()

```

# PCA

PCA on monthly log returns.

```{r pca1}

data_rets_m %>%
  filter(type != "index") %>% 
  ungroup() %>% 
  select(date, name, return_ln) %>% 
  # fill to remove the misssing values
  spread(name, return_ln, fill = 0) %>% 
  select(-date) %>% 
  as.matrix() %>% 
  prcomp(scale. = T, center = T) -> pca

pca %>% summary()

pca %>% fviz_eig()
pca %>% fviz_pca_biplot()

```

PCA loadings (eigenvectors) and eigenvalues.

```{r eigen}

pca$rotation

pca$sdev^2

```

# Varimax

Varimax rotation.

```{r varimax}

varimaxrot <- pca$rotation %>% varimax()

varimaxrot

```

For taking into account the maximum of the first three components, it seems that a combination of Bodycote, Hill&Smith, and Weir could be fine.

In a linear model.

```{r linmodel2}

lm(`FTSE 350 Industrial Engineering` ~ `Bodycote plc` + `Hill & Smith Holdings PLC` + `The Weir Group PLC` -1, 
   data = returns_daily %>% select(-date)) -> model_pca_daily

tidy(model_pca_daily)
# R-square
glance(model_pca_daily)$r.squared



lm(`FTSE 350 Industrial Engineering` ~ `Bodycote plc` + `Hill & Smith Holdings PLC` + `The Weir Group PLC` -1, 
   data = returns_monthly %>% select(-date)) -> model_pca_monthly

tidy(model_pca_monthly)
# R-square
glance(model_pca_monthly)$r.squared
```

Best fit portfolio return. Look at monthly here.

```{r bestfit}

returns_monthly %>% 
  mutate(fitted = model_pca_monthly$fitted.values) %>% 
  select(date, fitted, everything()) -> return_fitted

return_fitted %>% 
  ggplot(aes(x = `FTSE 350 Industrial Engineering`, y = fitted)) +
  geom_point(aes(color = fitted > `FTSE 350 Industrial Engineering`)) +
  geom_abline(slope=1, intercept=0) +
  geom_smooth(method = "lm") +
  theme_tq()


```

Cumulative return for performance.

```{r cumretfitted, warning=F}

data_rets_m_fitted <- return_fitted %>%
  gather(name, return_ln, -date) %>% 
  mutate(type = case_when(name == "fitted"        ~ "fitted",
                          str_detect(name, "FTSE") ~ "index",
                          T ~ "stock")) %>% 
  arrange(type, name, date) %>% 
  group_by(name, type) %>%
  mutate(cumret_ln = cumsum(return_ln),
         adjusted_ln = exp(cumret_ln))

data_rets_m_fitted %>% 
  ggplot(aes(x = date, y = adjusted_ln, color = name)) +
  geom_line(aes(size = factor(desc(type)), alpha = factor(desc(type)))) +
  scale_size_discrete(range = c(0.4,1.4), guide = "none") +
  scale_alpha_discrete(range = c(0.4, 1),  guide = "none") +
  theme_tq()

```

Tracking error etc. could be computed with `tidyquant::tq_performance()`, see https://cran.r-project.org/web/packages/tidyquant/vignettes/TQ05-performance-analysis-with-tidyquant.html