---
slug: "r-problems"
date: "2019-10-04"
title: "Top annoying R problems and how to overcome them"
tags: ["R"]
authors: [mgei]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, warning = F, message = F, cache = T)
```

This is to be extended each time I shall get accross one of my common issues.

# German Windows encoding and formatting trouble when reading data and sourcing other files

Problems: 

* files are not compatible between my (german) Windows machine and my Ubuntu system
* special cases like German Umlaute (ä-ö-ü) are not displayed properly on Windows
* reading in data with packages `readr` or `readxl` is troublesome

Solutions:

* save R or Rmd files in RStudio with encoding (*File > Save with Encoding...*)
    * use UTF-8
* read in data with encodinding
    * `readr::read_csv("character_list5.csv", locale = locale(encoding = 'latin1'))`
    * UTF-8
* if using MS Excel, save data CSV UTF-8
* `source()` other files with encoding: `source("file.R" encoding = "utf-8)`
    * This is especially crucial with Shiny where it's common to outsource functions or data reading to other files.
    
# Debugging a Shiny app

Problem:

You don't know what's happening in the background. Why is there no output? Has the data even be read in correctly?

Solution:

* use `print(data)`'s (or anything else that's stationary) and follow the output in the console
* use `reactive({ print(whatever())})` for Shiny reactive variables

# Early stopping and the use of return in R functions

Stop R session:

* `quit()`: terminates the R Session, RStudio will ask you to reboot after

Stop with warning:

* `stop("your warning message)`: end the script

Used to early terminate loops:

* `next()`: halts the current iteration and goes to the next
* `break()`: breaks out of a loop

Not used in R or other purpose:

* `exit()`
* `end()`: extracts observations of vectors or time series

Stop executing a function:

* `return(whatever)`

I learned that in R you do not need to specify `return()` in functions. It will throw out what's computed anyway. However, `return()` is great to stop the function from continuing to execute. For example, the following will not run if a non-NA input is given.

```{r returnfunction}

yolo <- function(x) {
  if (!is.na(x)) {
    return(x)
  }
  
  Sys.sleep(10)
  
  return("thanks for waiting, it's because you have not provided a non-NA input")
}

yolo("hello")

```

# Dplyr functions with column name as inputs

You want to write a function that has as the input the column name that is to be `dplyr::select`-ed or `mutate`-ed. Or even `summarise`-ed. It's not so trivial, because in standard dplyr code you do not need to put quotation marks, so the variables are not interpreted as such.

Here's how to do it:

```{r dplyrnames1}

library(dplyr)

select_and_summarise <- function(data, group_colname, mean_colname) {
  data %>%
    group_by(!!as.name(group_colname)) %>%
    summarise(mean = mean(sqrt(!!as.name(mean_colname))))
}

mydata <- as_tibble(mtcars)

select_and_summarise(mydata, "cyl", "mpg")

```

Now for `mutate` there's more needed:

```{r dplyrnames2}

example_mutate <- function(data, sqrt_colname, new_sqrt_colname){
  data %>%
    mutate(!!new_sqrt_colname := sqrt(!!as.name(sqrt_colname)))
}

example_mutate(mydata, "mpg", "mpg_sqrt")
```

And last but not least, what it's not just one colname but several. Say `c("mpg", "cyl", "hp")`?

```{r dplyrnames3}

select_several <- function(data, select_columns) {
  data %>% 
    select(!!paste0(lapply(select_columns, as.name)))
}

select_several(mydata, c("mpg", "cyl", "hp"))

```